simplex: An individual-based platform for ecological modeling

### 2015-10-26 paper/PrePrint.Rmd
modified PrePrint.Rmd and generated new PrePrint.pdf -KL

---
title: "simplex: An individual-based ecological modeling platform"
author: "Kenneth J. Locey,
Dept. of Biology, University of Indiana, Bloomington"
geometry: margin=2.54cm
output: pdf_document
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

>>>>

---
title: "simplex: An individual-based platform for ecological modeling"
author: "Kenneth J. Locey"
fontsize: 12pt
geometry: margin=2.54cm
output: pdf_document
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

\begin{center} 
Department of Biology, University of Indiana, Bloomington
\end{center}

\begin{center}
email: kjlocey@indiana.edu
\end{center}

\pagebreak

# Abstract

A popular approach in ecological modeling explicitly simulates the lives of individual organisms. Individual-based modeling (IBM) allows ecological complexity to be studied as it emerges from the individual-level, to explicitly synthesize theories and principles, and to track and record virtually any aspect of a system. I developed a novel IBM platform called simplex to explicitly simulate physiology, life history, resource-limited growth, fluctuating environmental dynamics, and community assembly and structure in both fluid and non-fluid environments. simplex accomplishes three primary tasks in keeping with its namesake (i.e., the generalized notion of a triangle to arbitrary dimensions). First, simplex assembles and runs stochastic ecological models from random combinations of over 20 state variables. Second, simplex stores the output of models and generates and stores animations. Third, simplex provides R Markdown files for analyzing simulated data. The simplex source code is freely available, includes the standard ODD description protocol, is open to community-based revision, includes unit testing and time testing, and has been developed to be run using modest hardware and freely available open-source tools.

\pagebreak
